test

### 2020-07-16 manuscript/response_to_reviews.tex

\textbf{Our preference is to keep this figure in the main manuscript. We
agree that similar patterns are common to any forest, and are not trying
to claim this as any kind of novel result, but we're not aware of
studies showing similar height profiles for a temperate forest. Such
patterns are not shown some relevant textbooks: Gordan Bonan's 2016
\emph{Ecological Climatology} and Norman \& Campbell's
\emph{Environmental Biophysics}. If we're missing studies that show such
vertical profiles, we'd appreciate the references.}

\textbf{We did create two new figures showing the results of the
tree-ring analysis (and moved two tables to the SI).}

\begin{itemize}
\tightlist
\item
  Line 330-331: any evidence to support this? It looks rather
  speculative. As suggested above, does any species particularly show
  positive Rt or positive Rt are randomly distributed in your trees?
\end{itemize}

\emph{Current text, with phrase of concern in bold: ``Across all three
droughts, the majority of trees experienced reduced growth, but a
substantial portion had increased growth (Fig. 1b), potentially due to
decreased leaf area of competitors during the drought (\textbf{REF--if
we can find one}), and consistent with prior observations that smaller
trees can exhibit increased growth rates during drought (Bennett
\emph{et al.}, 2015).'' }

\textbf{To address the question about species, we have added a new
figure (\emph{Fig. 2}) showing how \(Rt\) varied across species in each
of the three droughts.}

\begin{itemize}
\tightlist
\item
  Line 331-333: not so sure about this. It can also mean that there is
  no causal relationship between Rt and covariates, mostly because
  radial growth Rt should not be used as an integral proxy of tree
  resistance to drought, such as it is often used now in the literature.
\end{itemize}

\emph{Tree-ring data inherently have a lot of variation. Neil could
probably write a better response here, with refs.}

\begin{itemize}
\tightlist
\item
  Line 334-339: indeed, assessment of these two variables is not
  conclusive, as suggested above (just data from 2018, from the past
  estimated and correlated).
\end{itemize}

\textbf{Please see response above.}

\begin{itemize}
\tightlist
\item
  Line 348: non-drought.
\end{itemize}

\textbf{Fixed.}

\begin{itemize}
\tightlist
\item
  Line 349: do authors really think that lower radial growth as from Rt
  on a specific year means that the tree is more vulnerable? I think
  this is over conclusive.
\end{itemize}

\textbf{We did not intend the meaning here as vulnerability \emph{to
mortality} (assuming that's the objection here). We have reworded this
sentence to read, ``Belowground, taller trees would tend to have larger
root systems, but the potentially greater access to water did not
override the disadvantage conferred by height--and, in fact, greater
moisture access in non-drought years (here, higher TWI) appears to make
trees more sensitive to drought (Zuleta \emph{et al.}, 2017; Stovall
\emph{et al.}, 2019).''}

\begin{itemize}
\tightlist
\item
  Line 400: I do not agree we can directly link Rt with tree
  vulnerability to drought; we might be currently overinterpreting these
  resilience indices extrapolating to physiological conclusions that do
  not necessary need to hold.
\end{itemize}

\textbf{See above comment. We have reworded this sentence to read, ``In
the meantime, the results of this study advance our knowledge of the
factors conferring drought resistance in a mature forest, opening the
door for more accurate forecasting of forest responses to future
drought.''}

Table 1: do `-` signs in the table correspond to `no' (non-significant)?
Please explain.

\textbf{This is now clarified in the table footnote.}

@@

Table 3: please, add also the number of trees, it is more informative
than the number of cores to check sample size in your study.

\textbf{In this study, we have one core per tree, now specified in the
methods. We have changed the table header to ``n trees''.}

@@

Table 3: the authors could show some test to see if the traits presented
are different among species.

\textbf{We now show differences in traits among species in Table S4.}

\emph{need to add ANOVA, prsent reults in paper}

@@

Table S1 and S2: correct R2.

\emph{IM I added the correct Latex notation but it's not rendering it
when I knit the file\ldots{} (KAT: I just tried and wasn't able to get
it working.}

Table S3: please, explain what `rank' means.

\textbf{Done.}

@@

Figure S3: I understand that authors only have height data from 2018 and
height for the other dates looks trivial: indeed there will be an
increase in height with an increase in DBH in time. The question is
whether there was any stand dynamics involved in those height series
that had to be included in the models.

\textbf{We recognize the importance of stand dynamics, and have added
specific mention to the discussion: ``It is likely because of the
moderate impact of these droughts, along with other factors influencing
tree growth (e.g., stand dynamics), that our best models characterize
only a modest amount of variation,\ldots{}''}

\hypertarget{referee-2}{%
\subsection{Referee: 2}\label{referee-2}}

@@

The issue of the preferential death of large trees is not only important
because of its effects on forest function and culture. It is important
because it seems to require novel theory. Under current expectations,
the largest individuals, with their greater root reach, etc., should be
more rather than less resistant. That they might not be demands a novel
explanatory framework. Yet the very existence of the tendency is
debated, with some studies reporting greater mortality among small
rather than large individuals (Camarero et al., 2015; Colangelo et al.,
2017). These studies, however, were not designed to take into account
the microsite variation that presumably should lead to marked
differences in survivorship even within the same height class with
drought. One approach, which we have used (Soriano et al., in press;
Olson et al., 2018) is to conduct experiments under controlled
conditions to detect the effect of height, and our results are
consistent with the greater vulnerability of taller individuals.
However, under natural conditions, a study that takes this variation
into account is needed to test the notion that taller individuals are
preferentially vulnerable is needed, and this study provides exactly
such a design.

I found this paper not only to address a very important point but also
to be exceptionally well structured and written. My main suggestion is
that it could use a bit more emphasis on its importance, and I would be
interested in hearing a bit more about what the authors feel their
results imply regarding the cause of the preferential vulnerability of
large individuals. I give line-by-line suggestions (and sometimes just
reactions):

\textbf{Thank you for this wonderful review.}

\emph{What a nice review! We can easily add some language on the
importance. We should also add citations to many of the studies
mentioned.}

first paragraph---this setup is very compelling, and the point is
urgently true.

\textbf{Thank you}

66- ``in their basal portions'' seems fine for this level of intro, but
note that in angiosperms, taller trees also have wider vessels at the
stem tip (and thus throughout the conductive system) (Zach et al., 2010;
Olson et al., 2014, 2018). This seems plausibly the result of selection
keeping conduit carbon costs down with height growth. As a tree grows
taller, if terminal twig vessel diameters become wider, then given a
constant rate of tip-to-base vessel widening, vessel diameters will be
wider along the entire stem. Vessels that are wider throughout would
have lower per-conduit resistance than those that, all else being equal,
are narrower. So, conceivably as a tree gets taller, if vessels widen at
the twig tip, then lower per-conduit resistance could mean that the same
leaf area is supplied by fewer vessels. This could potentially
contribute to keeping per-leaf area carbon costs constant with height
growth. This prediction seems confirmed by the experiment of (Echeverría
et al., 2019), who found that as Moringa trees grew taller, there were
wider vessels at the stem tip and fewer vessels per unit leaf area, but
that conductance per unit leaf area should remain the same. Terminal
twig tracheids don't seem to widen in gymnosperms, e.g.~(Williams et
al., 2019). (Sack et al., 2012) have shown very clearly that petiole
base vessel diameter scales predictably with leaf size, as would be
expected; it is not clear how leaf size variation is involved in
terminal twig vessel diameter scaling with height, though in wide
samplings leaf size doesn't seem to scale with height (Jensen \&
Zwieniecki, 2013).

\textbf{Thank you for this interesting information. We've modified the
sentence to read, "Indeed, tall trees require xylem of greater hydraulic
efficiency, such that xylem conduit diameters are wider in the basal
portions of taller trees, both within and across species (Olson \emph{et
al.}, 2018; Liu \emph{et al.}, 2019), and throughout the conductive
systems of angiosperms (Zak et al.~2010, Olson et al.~2014,2018).}

67 Appealing to the vulnerability-diameter link might get blowback from
some readers because there is as yet no satisfactory theory linking
vulnerability to diameter in an entirely satisfactory way, so maybe just
say ``wider conduits are likely'' or that they are ``plausibly'' or
``possibly'' more vulnerable, or that many ``data are consistent'' with
such a link. My view is that there very well might not be a link, but in
the absence of a link between vulnerability and conduit diameter, it is
very hard to explain a very large number of well-established
observations. These include the following: why growth rings always
proceed from wide to narrow conduits as drought ensues; why maximum
community conduit diameter is narrower in drylands than in wet places;
why lianas have wider vessel diameter variances (i.e.~not only wider but
also numerous narrower vessels) than self-supporting plants (Rosell \&
Olson, 2014); why plants virtually always produce vessels narrower than
the maximum possible; when internal helical sculpture in conduits varies
in a growth ring it is always associated with narrow latewood vessels;
when simple and scalariform plates vary across growth rings sclariform
plates are always found in the narrow latewood vessels; and many other
observations. If there is no link between drought induced embolism
vulnerability and conduit diameter, and there might not be, all of these
observations, which are currently simultaneously and parsimoniously
explained by the link, will require alternative explanations. As a
result, while there is as yet no well accepted mechanistic link between
vulnerability and diameter, it seems to me irresponsible to reject the
link without providing a satisfactory and similarly parsimonious and
general replacement explanation for all of these phenomena. So, I think
the authors are justified in cautiously invoking the link here.

\textbf{We've added the word ``plausibly''. We agree that this linkage
makes a lot of sense.}

69 this is a very important point---microsite is crucial both for the
ultimate height an individual grows to as well as its vulnerability. The
generalization that on a global level large trees seem
disproportionately vulnerable is often greeted with the skeptical
observation that in specific cases a small tree died next to a large one
that survived. However, the prediction is that \emph{all else being
equal} large individuals should be more vulnerable than smaller
\emph{conspecifics}. There is so much variation in microsite and other
factors that the greater vulnerability of larger individuals should
emerge clearly only in carefully controlled experiments where all else
is indeed (virtually) equal, or with high sample sizes where variation
in microsite etc should random out and the signal of generally higher
vulnerability should come through.

\textbf{Thank you.}

73 The ``greater root reach'' point is almost universally appealed
to---cites for these counterarguments would be useful.

@@

73 Also, larger trees would necessarily seem to have greater stem
storage capacity and thus, if capacitance really is important, more
water volume available for buffering transient drought events, so this
is another reason to expect larger individuals to be more rather than
less resistant.

77 In the concluding sentence, ``greater growth reductions'' came as a
surprise because I thought that the focus was greater vulnerability/
mortality. Also, there are convincing theoretical and empirical reasons
to think that large trees don't slow down e.g.~(Sillett et al., 2010;
Selaya \& Anten, 2010; Michaletz et al., 2014).

@@

96 this all seems compelling. To make the point stronger, either justify
that the leaf traits are causally more important, or that, while they
make up part of a unified hydraulic system all of whose components are
important, leaf traits are simply easier to measure and this justifies
their study. The latter position seems most plausible to me.

102 great questions, though within-species comparisons are likely to be
most informative here, and your mixed model design is in accord with
this; just being cautious because sometimes it's seductive to regard
selection acting between rather than within species

\textbf{Thank you.}

105 this sounds great, but was a little unclear. When you say ``first,
we test\ldots{}'' I thought that you would examine a relationship
between microsite and some tree characteristic, e.g.~that the largest
trees are in moist microsites. But the next sentence didn't refer to a
tree-microsite correlation but vulnerability-size. It would be good to
clarify the prose here because these are very important passages. The
third sentence did seem to be about tree-microsite correlations.

112 wasn't clear where the ``ring porous more drought resistant''
hypothesis came from. Global data show that ring porous species are
almost entirely restricted to the north temperate zone, especially
deciduous broadleaf forests (Wheeler et al., 2007). They are virtually
absent in drylands, especially warm tropical drylands. Species in
drylands are almost overwhelmingly diffuse porous; the high wood density
species in chaparrals and in some desert plants can be semi ring porous
(Olson et al., in press).

\textbf{This is a great point. The hypothesis refers to temperate
deciduous biome, which we now clarify in the sentence in question:}

\textbf{``Studies in temperate broadleaf forests have observed that
ring-porous species showing higher drought tolerance than diffuse-porous
species (Kannenberg \emph{et al.}, 2019; Friedrichs \emph{et al.}, 2009;
Elliott \emph{et al.}, 2015), but this distinction would not hold in the
global context (\emph{Wheeler et al.~2007, Olson et al.~2020}) and does
not resolve differences among the many species within each category.''}

\textbf{We dug into the ring/diffuse porous distinction more, and ended
up making a few substantive changes that relate to this theme: (1) We
added a new figure showing \(Rt\) across species (Fig. 2). This
illustrates that the two diffuse-porous species, \emph{Liriodendron
tulipifera} and \emph{Fagus grandifolia}, were at opposite ends of the
\(Rt\) spectrum. (2) We modified the criteria of selecting candidate
variables for the full models to include the criterion that directions
of responses must be consistent across droughts for inclusion in the
full model. This resulted in exclusion of xylem porosity (Table S4)--a
change that is also warrented by the fact that we only have two
diffuse-porous species. (3) We added the following content to the
discussion: }

\emph{insert text}

113 These correlations seem reasonable, e.g.~see (Méndez-Alonzo et al.,
2012)

124 and Table 3. This seems like a good palette of species because they
are very different in their hydraulic construction. In addition to
libriform fibers, oaks have very abundant vasicentric tracheids,
especially in latewood. This means that the oaks will likely maintain
some conductance even under highly negative tensions. This is not by
virtue of their ring porosity (note that most oaks globally aren't ring
porous!), but because of the high proportion of vasicentric tracheids in
their latewood. Liriodendron (diffuse porous, scalariform perforation
plates) and Carya (likely ring porous at your locality; simple
perforation plates) both have libriform fibers and lack vasicentric
tracheids. See the data in (Olson et al., in press). There is relatively
low family-level diversity represented here; our phylogenetic analyses
(Olson et al., in press) suggest that there is very little phylogenetic
signal in the conduit diameter-stem length relationship, so this could
help justify if necessary not including phylogeny in the analyses.

\emph{this is new terminology to me. Look into Olson et al.~in press.
Would these wood traits be worth including as covariates?}

127 ``tree size''= height?

@@

138 you might discuss the adequacy of the height range studied. If it is
true that conduit diameter predicts vulnerability, then the greatest
differences in vulnerability should be over the height range over which
conduit diameter changes most rapidly. The relationship between conduit
diameter and plant height turns out to follow a power-law like
distribution, see for example the SI in (Olson et al., 2018). This means
that differences in vulnerability should be very marked across smaller
size classes (say below 7 m or so and very marked below 3 m) and be less
and less visible across larger plants. So, studies of changes in conduit
diameter/ vulnerability with height would ideally include plants smaller
than those included here, so maybe comment on how the height range might
affect the ability to detect the patterns of interest. It strikes me
that it is appropriate because it is the height range over which the
putative preferential susceptibility of large trees to drought is being
observed.

173 TWI seems like an appropriate index for characterizing microsite,
with the expectation that individuals in moister areas should be more
buffered from drought.

\textbf{Thank you.}

209 It couldn't hurt to define basal area increment

227 the nesting seems appropriate and sufficient for taking species/
``phylogeny'' into account

@@

260 amazing

\textbf{Thank you.}

275---some people think that the increase in terminal twig vessels with
height growth keeps pace with the changes in VPD from understory to
canopy

@@

301 this intro paragraph to the discussion ranks very well the results
from most significant to least

\textbf{Thank you.}

327 I agree that this is the point that most deserves discussion in
paragraph 2

334 You might go into detail regarding what you think this says about
the causal hypothesis behind the global decline of large individuals. To
me it helps reject notions that somehow larger trees are more exposed to
wind throw or other damaging agents. It is consistent with the
hypothesis that larger individuals have wider conduits, and (assuming
that wider conduits are more vulnerable to drought induced embolism)
thus larger individuals are (all else being equal\ldots) more vulnerable
to embolism than smaller conspecifics.

336 beautifully said ``impossible for trees to e ciently transport water
to great heights and simultaneously maintain strong resistance and
resilience to drought-induced embolism''

\textbf{Thank you.}

@@

385 very nice paragraph

\textbf{Thank you.}

In conclusion, this ms addresses a very important issue with a
well-designed study and appropriate analyses. It is well written and
interesting and I hope my comments are useful.

\textbf{Thank you.}

Mark Olson
\href{mailto:molson@ib.unam.mx}{\nolinkurl{molson@ib.unam.mx}}

Camarero JJ, Gazol A, Sangüesa-Barreda G, Oliva J, Vicente-Serrano SM.
2015. To die or not to die: early warnings of tree dieback in response
to a severe drought (D Gibson, Ed.). Journal of Ecology 103: 44--57.

Christoffersen BO, Gloor M, Fauset S, Fyllas NM, Galbraith DR, Baker TR,
Kruijt B, Rowland L, Fisher RA, Binks OJ, et al.~2016. Linking hydraulic
traits to tropical forest function in a size-structured and trait-driven
model (TFS v.1-Hydro). Geoscientific Model Development 9: 4227--4255.

@@

Colangelo M, Camarero JJ, Borghetti M, Gazol A, Gentilesca T, Ripullone
F. 2017. Size Matters a Lot: Drought-Affected Italian Oaks Are Smaller
and Show Lower Growth Prior to Tree Death. Frontiers in Plant Science 8.

Echeverría A, Anfodillo T, Soriano D, Rosell JA, Olson ME. 2019.
Constant theoretical conductance, changes in vessel diameter and number
with height growth in Moringa oleifera. Journal of Experimental Botany.

Jensen KH, Zwieniecki MA. 2013. Physical Limits to Leaf Size in Tall
Trees. Physical Review Letters 110.

Méndez-Alonzo R, Paz H, Zuluaga RC, Rosell JA, Olson ME. 2012.
Coordinated evolution of leaf and stem economics in tropical dry forest
trees. Ecology 93: 2397--2406.

Michaletz ST, Cheng D, Kerkhoff AJ, Enquist BJ. 2014. Convergence of
terrestrial plant production across global climate gradients. Nature
512: 39--43.

Olson ME, Anfodillo T, Rosell JA, Petit G, Crivellaro A, Isnard S,
León-Gómez C, Alvarado-Cárdenas LO, Castorena M. 2014. Universal
hydraulics of the flowering plants: vessel diameter scales with stem
length across angiosperm lineages, habits and climates. Ecology Letters
17: 988--997.
